Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
results

As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that a subject was in. The predictive performance of Naive Bayes, Linear SVM, and Multivariate was compared. We observed that different feature selection and extraction methods, as well as different structure on the data affects the predictive performance of the classifiers.

Table~\ref{tab:results} illustrates the overall performance of the three classifiers. Rows indicate various feature selection methods. We applied PCA to the original data, while the other classifiers were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}). For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation. Columns depict the used classifiers. For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.

Furthermore, to test the significance of the differences in performance between the different methods we performed an exact binomial test in each case, from which we obtained the confidence interval. The results were rounded to one decimal point. For the following comparison we are only looking at the case where \emph{Norm/Vox-M-SD ROIs} was used.

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used. The confidence interval of the predictive performance of Naive Bayes when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...

Comparing classifiers when the same data structure is used, we can distinguish the fact that Multivariate performs better than Naive Bayes. Also that Linear SVM performs better than Naive Bayes, though not significantly. The confidence interval of Naive Bayes is 95\% $CI: 68.9-77.0$ and of Linear SVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.

Moreover, when different data structure is used the performance increases, though not significantly. This can be seen by comparing the performance of Naive Bayes when different data structure is used. In the first case it is 73.13\% while in the second case is 76.88\%. The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.

Finally, we can distinguish the fact that Multivariate outperforms all the other classifiers, though we did not try a different data structure in the case of Linear SVM. However, we still do not expect that the performance of Linear SVM will be higher that this of Multivariate. The confidence interval of Multivariate is 95\% $CI: 75.7-83.1$.

>>>>

As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that a subject was in. The predictive performance of GNB, LSVM, and MVG was compared. We observed that different feature selection and extraction methods, as well as different structure on the data affects the predictive performance of the classifiers.

Table~\ref{tab:results} illustrates the overall performance of the three classifiers. Rows indicate various feature selection methods. We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}). For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation. Columns depict the used classifiers. For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.

Furthermore, to test the significance of the differences in performance between the different methods we performed an exact binomial test in each case, from which we obtained the confidence interval. The results were rounded to one decimal point. For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...

Comparing classifiers when the same data structure is used, we can distinguish the fact that MVG performs better than GNB. Also that LSVM performs better than GNB, though not significantly. The confidence interval of GNB is 95\% $CI: 68.9-77.0$ and of LSVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.

Moreover, when different method for structuring the data is used the performance increases, though not significantly. This can be seen by comparing the performance of GNB when different method for structuring the data is used. In the first case it is 73.13\% while in the second case is 76.88\%. The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.

Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different data structure in the case of LSVM. However, we still do not expect that the performance of LSVM will be higher that this of MVG. The confidence interval of MVG is 95\% $CI: 75.7-83.1$.
